---
title: "Analysis"
author: "William Procter"
date: "`r Sys.Date()`"
output: html_document
editor_options:
  markdown:
    wrap: sentence
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../../docs/report") })
---


# Abstract

Accurate assessment of risk is an essential for effective response to any natural disaster. The methodologies
used to assess risk can end up underestimating vulnerabilities. Tropical Storm Irene offers an example of
inadequate assessment of risk, which then leads to inadequate planning for and response to a disaster. The
storm inundated Vermont with unprecedented rainfall on August 28 and 29 of 2011. The storm destroyed
480 bridges and 960 culverts (where streams cross under a road), causing $350 million in road damage and
cutting off road access to 13 mountain communities. Even Vermont's emergency management offices were
flooded! Some of the most affected people were living in mobile homes, whether on individual parcels of
land or in mobile home parks. At least 130 mobile homes were destroyed and an additional 300 severely
damaged (Figure 1). Our problem will evaluate assessments of flooding risks with a focus on mobile homes
in Vermont.

There are two different ways of assessing flooding risk in Vermont: one is by the federal agency, FEMA
(The Federal Emergency Management Association), and one by a state agency, Vermont Rivers Program.
The federal agency, FEMA, estimates flood risk in terms of inundation from rising water levels in stable
river channels. Based on existing channels, FEMA hydrologists estimate the region of land that would be
potentially flooded by a 1% (100-year) flood. The residents with mortgages in that region are required to
purchase flood insurance.

The state of Vermont’s River Corridors Program estimates flooding risk differently, using river corridors.
After Irene, the state of Vermont recognized that the most damaging flooding in Vermont is not due to
inundation but rather due to fluvial erosion: the erosion of riverbanks as the river channel widens or
migrates to form new channels (Figure 1 and Figure 2). By this estimation, regions where rivers may erode
and migrate to in the future are also at risk of flooding.

The research involved a spatial overlay analysis using existing and
newly created data sets to assess the vulnerability of mobile home
parks to flooding (Fig. 3). The analysis used a number of publicly
available data sets, including Vermont E911, a mobile home park
registry, FIRMs, and orthographic imagery.


This is a reproduction of a study that is taught in Middlebury College's Human Geography with GIS course, which itself builds on Baker et al's 2011 study on **Rapid Flood Exposure Assessment of Vermont Mobile Home Parks Following Tropical Storm Irene**

Write a brief abstract about your research project.

If the project is a reproduction or replication study, include a declaration of the study type with a full reference to the original study.
For example:

This study is a *replication* of:

> citation to prior study

A graphical abstract of the study could also be included as an image here.

# Study metadata

- `Key words`: Comma-separated list of keywords (tags) for searchability. Geographers often use one or two keywords each for: theory, geographic context, and methods.
- `Subject`: select from the [BePress Taxonomy](http://digitalcommons.bepress.com/cgi/viewcontent.cgi?article=1008&context=reference)
- `Date created`: date when project was started
- `Date modified`: date of most recent revision
- `Spatial Coverage`: Specify the geographic extent of your study. This may be a place name and link to a feature in a gazetteer like GeoNames or OpenStreetMap, or a well known text (WKT) representation of a bounding box.
- `Spatial Resolution`: Specify the spatial resolution as a scale factor, description of the level of detail of each unit of observation (including administrative level of administrative areas), and/or or distance of a raster GRID size
- `Spatial Reference System`: Specify the geographic or projected coordinate system for the study, e.g. EPSG:4326
- `Temporal Coverage`: Specify the temporal extent of your study---i.e. the range of time represented by the data observations.
- `Temporal Resolution`: Specify the temporal resolution of your study---i.e. the duration of time for which each observation represents or the revisit period for repeated observations
- `Funding Name`: name of funding for the project
- `Funding Title`: title of project grant
- `Award info URI`: web address for award information
- `Award number`: award number

## Original study spatio-temporal metadata

- `Spatial Coverage`: extent of original study
- `Spatial Resolution`: resolution of original study
- `Spatial Reference System`: spatial reference system of original study
- `Temporal Coverage`: temporal extent of original study
- `Temporal Resolution`: temporal resolution of original study

# Study design

Describe how the study relates to prior literature, **reproduction study**, **reanalysis study**, or **replication study**?

Also describe the original study archetype, e.g. is it **observational**, **experimental**, **quasi-experimental**, or **exploratory**?  Original study is observational

Enumerate specific **hypotheses** to be tested or **research questions** to be investigated here, and specify the type of method, statistical test or model to be used on the hypothesis or question.

# Materials and procedure

## Computational environment

```{r environment-setup, include = FALSE}
# record all the packages you are using here
# this includes any calls to library(), require(),
# and double colons such as here::i_am()
packages <- c("downloader", "rdhs", "haven", "readr", "tidyverse", "pastecs",
             "stars", "sf", "sp", "classInt", 
             "tmap", "units", "knitr", "tibble", "here", "dplyr", "rgdal")

# force all conflicts to become errors
# if you load dplyr and use filter(), R has to guess whether you mean dplyr::filter() or stats::filter()
# the conflicted package forces you to be explicit about this
# disable at your own peril
# https://conflicted.r-lib.org/
require(conflicted)

# load and install required packages
# https://groundhogr.com/
if (!require(groundhog)) {
  install.packages("groundhog")
  require(groundhog)
}

# this date will be used to determine the versions of R and your packages
# it is best practice to keep R and its packages up to date
groundhog.day <- "2023-11-27"
set.groundhog.folder("../../data/scratch/groundhog/")

# this replaces any library() or require() calls
groundhog.library(packages, groundhog.day)
# you may need to install a correct version of R
# you may need to respond OK in the console to permit groundhog to install packages
# you may need to restart R and rerun this code to load installed packages
# In RStudio, restart r with Session -> Restart Session

# record the R processing environment
# alternatively, use devtools::session_info() for better results
writeLines(
  capture.output(sessionInfo()),
  here("procedure", "environment", paste0("r-environment-", Sys.Date(), ".txt"))
)

# save package citations
knitr::write_bib(c(packages, "base"), file = here("software.bib"))

# set up default knitr parameters
# https://yihui.org/knitr/options/
knitr::opts_chunk$set(
  echo = FALSE, # Run code, show outputs (don't show code)
  fig.retina = 4,
  fig.width = 8,
  fig.path = paste0(here("results", "figures"), "/")
)
```


## Set up file path shortcuts
```{r}
# these values allow you to access private and public raw data more efficiently
private_r <- here("data", "raw", "private")
public_r <- here("data", "raw", "public")
public_d <- here("data", "derived", "public")
results_figures <- here("results", "figures")
results_tables <- here("results", "tables")
scratch <- here("data", "scratch")
```



## Data and variables

There are four layers for this analysis, each coming from a different primary source.  Primary data sources for the study are to include ...

1. e911pts.shp - point - epsg: 32145
e911 point location data for all residences and buildings in Vermont, for use with emergency response.  The data file can be found on the Vermont Open GeoData Portal <http://geodata.vermont.gov/>
- SITETYPE: type of building structure/use case of the structure.  "MOBILEHOME" is the SITETYPE that indicates a site is a mobile home/

2. FEMA_100yr.shp - polygon - epsg: 32145
FEMA Flood Zone polygons with codes.  Codes starting with "A" indicate a 100-year flood risk zone. The data file can be found on the Vermont Open GeoData Portal <http://geodata.vermont.gov/>
- FLD_Zone: contains FEMA Flood Zone Codes.  If a polygon has a code begins with an "A", then that polygon indicates a 100-year flood zone.

3. river_corridors.shp - polygon - epsg: 32145
Vermont river corridor polygons, as defined by Flood Ready Vermont.  This flood hazard approach includes streams (with a 50 foot buffer) and rivers with watersheds more than 2km.  The data file can be found on the Vermont Open GeoData Portal <http://geodata.vermont.gov/>
- 

4. block_groups.shp - polygon - epsg: 32145
Census block group polygons in southern Vermont, with data on housing.  The data file was acquired from the US Census ACS Survey 2014-2018
- mobileHU: estimated total number of mobile home housing units within the block group
- totalHU: estimated total number of all housing units within the block group
- county: name of county in which the block_group is located


Each of the next subsections describes one data source.

### Primary data source1 name

- `Title`: Title of data source
- `Abstract`: Brief description of the data source
- `Spatial Coverage`: Specify the geographic extent of your study. This may be a place name and link to a feature in a gazetteer like GeoNames or OpenStreetMap, or a well known text (WKT) representation of a bounding box.
- `Spatial Resolution`: Specify the spatial resolution as a scale factor, description of the level of detail of each unit of observation (including administrative level of administrative areas), and/or or distance of a raster GRID size
- `Spatial Reference System`: Specify the geographic or projected coordinate system for the study
- `Temporal Coverage`: Specify the temporal extent of your study---i.e. the range of time represented by the data observations.
- `Temporal Resolution`: Specify the temporal resolution of your study---i.e. the duration of time for which each observation represents or the revisit period for repeated observations
- `Lineage`: Describe and/or cite data sources and/or methodological steps planned to create this data source.
  - sampling scheme, including spatial sampling
  - target sample size and method for determining sample size
  - stopping criteria for data collection and sampling (e.g. sample size, time elapsed)
  - de-identification / anonymization
  - experimental manipulation
- `Distribution`: Describe who will make the data available and how?
- `Constraints`: Legal constraints for *access* or *use* to protect *privacy* or *intellectual property rights*
- `Data Quality`: State any planned quality assessment
- `Variables`: For each variable, enter the following information. If you have two or more variables per data source, you may want to present this information in table form (shown below)
  - `Label`: variable name as used in the data or code
  - `Alias`: intuitive natural language name
  - `Definition`: Short description or definition of the variable. Include measurement units in description.
  - `Type`: data type, e.g. character string, integer, real
  - `Accuracy`: e.g. uncertainty of measurements
  - `Domain`: Expected range of Maximum and Minimum of numerical data, or codes or categories of nominal data, or reference to a standard codebook
  - `Missing Data Value(s)`: Values used to represent missing data and frequency of missing data observations
  - `Missing Data Frequency`: Frequency of missing data observations: not yet known for data to be collected

| Label | Alias | Definition | Type | Accuracy | Domain | Missing Data Value(s) | Missing Data Frequency |
| :--: | :--: | :--: | :--: | :--: | :--: | :--: | :--: |
| variable1 | ... | ... | ... | ... | ... | ... | ... |
| variable2 | ... | ... | ... | ... | ... | ... | ... |

### Primary data source2 name

... same form as above...




## Prior observations  

Prior experience with the study area, prior data collection, or prior observation of the data can compromise the validity of a study, e.g. through p-hacking.
Therefore, disclose any prior experience or observations at the time of study pre-registration here, with example text below:

At the time of this study pre-registration, the authors had _____ prior knowledge of the geography of the study region with regards to the ____ phenomena to be studied.
This study is related to ____ prior studies by the authors

For each primary data source, declare the extent to which authors had already engaged with the data:

- [ ] no data collection has started
- [ ] pilot test data has been collected
- [ ] data collection is in progress and data has not been observed
- [ ] data collection is in progress and __% of data has been observed
- [ ] data collection is complete and data has been observed. Explain how authors have already manipulated / explored the data.

For each secondary source, declare the extent to which authors had already engaged with the data:

- [ ] data is not available yet
- [ ] data is available, but only metadata has been observed
- [ ] metadata and descriptive statistics have been observed
- [ ] metadata and a pilot test subset or sample of the full dataset have been observed
- [ ] the full dataset has been observed. Explain how authors have already manipulated / explored the data.

If pilot test data has been collected or acquired, describe how the researchers observed and analyzed the pilot test, and the extent to which the pilot test influenced the research design.

## Bias and threats to validity

Given the research design and primary data to be collected and/or secondary data to be used, discuss common threats to validity and the approach to mitigating those threats, with an emphasis on geographic threats to validity.

These include:
  - uneven primary data collection due to geographic inaccessibility or other constraints
  - multiple hypothesis testing
  - edge or boundary effects
  - the modifiable areal unit problem
  - nonstationarity
  - spatial dependence or autocorrelation
  - temporal dependence or autocorrelation
  - spatial scale dependency
  - spatial anisotropies
  - confusion of spatial and a-spatial causation
  - ecological fallacy
  - uncertainty e.g. from spatial disaggregation, anonymization, differential privacy
  
MAUP

## Data transformations

Describe all data transformations planned to prepare data sources for analysis.
This section should explain with the fullest detail possible how to transform data from the **raw** state at the time of acquisition or observation, to the pre-processed **derived** state ready for the main analysis.
Including steps to check and mitigate sources of **bias** and **threats to validity**.
The method may anticipate **contingencies**, e.g. tests for normality and alternative decisions to make based on the results of the test.
More specifically, all the **geographic** and **variable** transformations required to prepare input data as described in the data and variables section above to match the study's spatio-temporal characteristics as described in the study metadata and study design sections.
Visual workflow diagrams may help communicate the methodology in this section.

Examples of **geographic** transformations include coordinate system transformations, aggregation, disaggregation, spatial interpolation, distance calculations, zonal statistics, etc.

Examples of **variable** transformations include standardization, normalization, constructed variables, imputation, classification, etc.

Be sure to include any steps planned to **exclude** observations with *missing* or *outlier* data, to **group** observations by *attribute* or *geographic* criteria, or to **impute** missing data or apply spatial or temporal **interpolation**.


# Read in the layers
```{r}
e911 <- read_sf(here(public_d, "e911pts.shp"))

FEMA_100yr <- read_sf(here(public_d, "FEMA_100yr.shp"))
  
river_corridors <- read_sf(here(public_d, "river_corridors.shp"))

block_groups <- read_sf(here(public_d, "block_groups.shp"))

towns <- read_sf(here(public_d, "towns.shp"))

ct_river_corridor <- read_sf(here(public_d, "CT_corridor_final_dissolved_buffered.shp"))
```


# Exploratory Data Analysis

## Visualize the FEMA flood zones and river corridors
```{r}
# FEMA flood zones
tmap_mode("plot")
fz_map <- tm_shape(block_groups) +
  tm_fill(col="tan", alpha=0.7, title="") +
  tm_shape(FEMA_100yr) +
  tm_fill(col="blue", alpha=0.7, legend.show = TRUE, title="FEMA Flood Zones") +
  tm_add_legend(title = "Flood Zones") +
  tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = 1, asp=0.8)
tmap_save(fz_map, filename = paste0(results_figures, "/FEMA_flood_zone_map.pdf")) #export to results/figures





# River corridors
rc_map <- tm_shape(block_groups) +
  tm_fill(col="tan", alpha=0.7, title="") +
  tm_shape(river_corridors) +
  tm_fill(col="darkblue", alpha=0.7, legend.show = TRUE, title="FEMA Flood Zones") +
  tm_add_legend(title = "River Corridors") +
  tm_layout(legend.width=3, legend.text.size = .6, legend.title.size = 1, asp=0.8)
tmap_save(rc_map, filename = paste0(results_figures, "/river_corridor_map.pdf")) #export to results/figures

fz_map
rc_map
```



# Clean up the flood zone and river corridor layers into more usable formats

## Clean up the FEMA data to identify all the zones that start with "A" (aka in the 100 year flood zone)...treat them all the same
```{r}
# Create a column called "flood" that is TRUE for all flood zones that are 100 year flood zones
FEMA_100yr <- FEMA_100yr %>%
  mutate(flood = case_when(FLD_ZONE == "A" ~ TRUE,
                           FLD_ZONE == "AE" ~ TRUE,
                           FLD_ZONE == "AO" ~ TRUE,
                           TRUE ~ FALSE))

table(FEMA_100yr$FLD_ZONE)
FEMA_100yr
```


## Group by flood (all of the 100 year flood zones) and dissolve the geometry to get a single multipart flood zone
```{r}
FEMA_multipart <- FEMA_100yr %>%
  group_by(flood) %>%
  summarize(geometry = st_union(geometry))
```


## And do the same thing for river corridors...create a single multipart river corridor 
```{r}
river_corridors_multipart <- river_corridors %>%
  mutate(flood = TRUE) %>%
  group_by(flood) %>%
  summarize(geometry = st_union(geometry))
```



# Calculating table columns to reproduce Table 1

## Column 1 (step 1): calculate the total number of mobile homes in each county from the ACS data - aggregate from the block groups
```{r}
# Group block groups by county, select just county and mobileHU variables, and for each county, sum the number of mobile homes from all of the block groups that lay within the county
MHs_by_county <- block_groups %>%
  group_by(county) %>%
  dplyr::select(county, mobileHU) %>%
  summarise(number_of_MHs = sum(mobileHU)) %>%
  st_drop_geometry()

MHs_by_county 
```



## Columns 3 and 4 (steps 3 and 4): mobile homes at risk

I exclude column 2, which was created using area weighted aggregation of mobile homes.However, the GEOG 0261 instructors discovered that this approach is less accurate than using the e911 point data to identify mobile homes at risk, so I will forego reproducing the AWR approach.


### Isolate only the points/residences that are mobile homes from the e911 data
```{r}
e911_mh <- e911 %>%
  dplyr::filter(sitetype == "MOBILE HOME")
```


### Add county and town variable (from the census/ACS data) to the mobile homes points, joining by location
```{r}
# Join attributes by location...identify the mobile home points that are spatially located within each county
mh_with_county <- st_join(e911_mh, block_groups, join = st_intersects)

# Also do this for towns
mh_with_county <- st_join(mh_with_county, towns, join = st_intersects)


# Group MHs by county and sum them for each county to get column 1...****WHEN I DID THIS THE COLUMN 1 WAS INCORRECT, INVESTIGATE THIS DEVIATION!!!
#MHs_by_county <- mh_with_county %>%
#  group_by(county) %>%
#  summarise(count_of_mobile_homes = n())
```

NOTE: I attempted also to sum the number of e911 mobile home points for each county.  This yielded a different number of mobile homes than Column 1 indicates.  Because the ACS measurement of `mobileHU` is a survey-based estimate, it does not represent the true number of mobile home structures.  This is a source of geographic uncertainty to this analysis, specifically an issue of spatial heterogeneity and construct validity.

I will proceed with using Total Number of Mobile Homes column from the ACS data to be consistent with the GEOG 0261 analysis, but this is something that may want to be changed in the future.



### Add buffer (18.3 meters aka 60 feet) to mobile home points to account for their structure sizes
```{r}
mh_buffer <- st_buffer(mh_with_county, dist = 18.3) #18.3 meters (60 feet) is the average length of a mobile home

# This does NOT dissolve the resulting geometries 
```


### Identify the mobile homes within FEMA flood zone, and group by county to get total number of MHs in the flood zone by county
```{r}
# Spatial join to the overall flood zone layer...not every MH will be within the flood zone
# Identifies which buffered mobile home points intersect flood zones
mh_fema_everything <- st_join(mh_buffer, FEMA_multipart, join = st_intersects)

# Filter to just the MHs that ended up in the flood zone...MHs in the flood zone meet the condition flood = TRUE
mh_fema <- mh_fema_everything %>%
  dplyr::filter(flood == TRUE)

# OR, just use a select by location, but this I found is more computationally intensive
mh_fema2 <- st_intersection(mh_buffer, FEMA_multipart)




# Group by county, counting the number of MHs at risk...using spatial join then filtered data
mh_fema_by_county <- mh_fema %>%
  group_by(county) %>%
  count(county, name = "mobile_home_count") %>%
  st_drop_geometry()

mh_fema_by_county


# Group by county, counting the number of MHs at risk...using select by location data
mh_fema_by_county2 <- mh_fema2 %>%
  group_by(county) %>%
  count(county, name = "mobile_home_count") %>%
  st_drop_geometry()

mh_fema_by_county2

# Both approaches seem to work!  But I get a warning for the select by location approach.  Thus, I will use the spatial join + filter data for use going forward.
```


### Now do the same thing but for river corridors...

### Identify the mobile homes within river corridors, and group by county to get total number of MHs in the river corridors by county
```{r}
# Spatial join to the overall river corridor layer...not every MH will be within the river corridors
# Identifies which buffered mobile home points intersect river corridors
mh_rc_everything <- st_join(mh_buffer, river_corridors_multipart, join = st_intersects)

# Filter to just the MHs that ended up in the flood zone...MHs in the river corridors meet the condition flood = TRUE
mh_rc <- mh_rc_everything %>%
  dplyr::filter(flood == TRUE)

# Group by county, counting the number of MHs at risk
mh_rc_by_county <- mh_rc %>%
  group_by(county) %>%
  count(county, name = "mobile_home_count") %>%
  st_drop_geometry()

mh_rc_by_county

# I use the spatial join + filter for the MHs in the river corridor approach due to the reasons outlined in the previous code chunk
```



## Join all columns together to get the final table!  Table 1:
```{r}
final_table_1 <- MHs_by_county %>%
  mutate(MHs_at_risk_FEMA = mh_fema_by_county$mobile_home_count) %>%
  mutate(MHs_at_risk_River_Corridors = mh_rc_by_county$mobile_home_count) %>%
  mutate(FEMA_rate = MHs_at_risk_FEMA / number_of_MHs) %>%
  mutate(RC_rate = MHs_at_risk_River_Corridors / number_of_MHs)
final_table

# Export the final table to a CSV
write.csv(final_table_1, file = file.path(results_tables, 'Table1.csv'), row.names = FALSE)
```

Unplanned deviation for reproduction: I decided to calculate "rates "risk rates" to indicate what proportion of a county's mobile homes lie within the FEMA flood zones and the River Corridors, respectively.  They are pretty similar for Windham County.  In Bennington County, the FEMA risk rate is higher than the River Corridor risk rate.  In Windsor County and Rutland County, the River Corridor risk rate is higher than the FEMA risk rate.




# Visualize MH flood risk by town:


## First, I need to find the number of mobile homes by town

```{r}
# Identify which mobile home points are in each town, and count up how many are in each town

mh_towns <- st_join(towns, e911_mh, join = st_intersects)

mh_towns_grouped <- mh_towns %>%
  group_by(townName) %>%
  count(townName, name = "mobile_home_count") %>%
  rename(town = townName)
```



## Next, I need to find which mobile homes are in EITHER a FEMA flood zone OR River Corridor

This will cast a wider net than if we look at just flood zones or river corridors individually, as this will maximize the number of mobile homes that are determined to be at risk.  At this stage in the analysis, we care more about seeing which towns have the highest vulnerability of mobile homes to flooding, not whether the VT River Corridor or FEMA Flood Zone approach is more accurate.  Thus, including both flood risk identification metrics is a safer approach to ensure we identify all mobile homes that are at some level of risk to flooding. 

```{r}
# Drop geometry from MHs identified to be in the FEMA flood zones
mh_fema_nogm <- mh_fema %>%
  st_drop_geometry()

# Drop geometry from MHs identified to be in the River Corridors
mh_rc_nogm <- mh_rc %>%
  st_drop_geometry()

# Full join the above two data frames...any mobile home that is in one or both flood hazard identifiers will get its own row in this data frame
mh_flooded <- full_join(mh_fema_nogm, mh_rc_nogm, by = "OBJECTID")

# Create a new column for town name that is the correct town name for where the mobile home is situated
mh_flooded <- mh_flooded %>%
  mutate(town = ifelse(!is.na(townName.x), townName.x, townName.y)) %>%
  select(-townName.x, -townName.y)

# Group by town, and count the number of at risk mobile homes in each town
mh_flooded_by_town <- mh_flooded %>%
  group_by(town) %>%
  count(town, name = "at_risk_count")
```

## Create a table to show the 10 towns with the highest mobile home flooding risk
```{r}
# Join number of mobile homes at risk to the total number of mobile homes in each town, by town
town_data <- full_join(mh_towns_grouped, mh_flooded_by_town, by = "town")

# Change N/As to zero...it would be N/A if there are no mobile homes at risk in that town
town_data <- town_data %>%
  mutate(at_risk_count = ifelse(is.na(at_risk_count), 0, at_risk_count))

# Calculate an at-risk rate
town_data <- town_data %>%
  mutate(pct_mh_at_risk = at_risk_count / mobile_home_count)

most_vulnerable_towns <- town_data %>%
  arrange(-pct_mh_at_risk) %>%
  head(10)

most_vulnerable_towns

# Export the table to a CSV
write.csv(most_vulnerable_towns, file = file.path(results_tables, 'TownsWithHighestPercentageOfMobileHomesAtRisk.csv'), row.names = FALSE)
```

Results slightly differ between this ant the GEOG 0261 results because I directly downloaded the towns.shp file from the VT Open GeoData Portal, while the GEOG 0261 class uses a pre-cleaned provided towns shapefile layer.  I could not use the provided one due to file corruption issues.  However, the provided one for the class distinguishes between Rutland Town and Rutland City, while the layer that I downloaded treated the two as a combined town of "Rutland." Rutland Town has a pct_mh_at_risk value of 57.89, so if my towns file distinguished just the town portion, it would be in the top 10 of highest risk towns for mobile homes at risk.


## Plot a choropleth map of the % of mobile homes in each town that are at risk to flooding, and export
```{r}
# Visualize!
tmap_mode("plot")
pct_mh_at_risk_by_town <- tm_shape(town_data) +
  tm_fill(col="pct_mh_at_risk", palette=rev(hcl.colors(4, "Purple-Blue")), style = "jenks", alpha=0.7, title="") +
  tm_text("town", size = 0.3) +
  tm_borders(lwd = 0.5, col = "black") +
  tm_layout(legend.width=3, legend.text.size = .6, title.size = .8, legend.title.size = .1, asp=1.7, title = "% of Mobile Homes at Risk By Town (Either FEMA or River Corridor)")

# Export map 
tmap_save(pct_mh_at_risk_by_town, filename = paste0(results_figures, "/pct_mh_at_risk_by_town.pdf")) #export to results/figures


pct_mh_at_risk_by_town
```

# Unplanned deviation for reproduction: identify areas where FEMA 100 year Flood Zones and VT River Corridors do not line up

I was curious about the discrepancy between the two metrics so created a layer that differences the two polygon layers and plots it with a satellite base-map.  Notably, the River Corridors include fewer lakes/ponds and large rivers.  This is significant, because these water bodies can still cause severe flooding if water influx causes them to over spill their banks.  Additionally, note that the Connecticut River is not included at all in the River Corridors.

```{r}
### Discrepancy between FEMA and RC
non_overlapping <- st_difference(FEMA_multipart, river_corridors_multipart)

# Interactive map so that the reader can investigate where the discrepancies occur!
tmap_mode("view")
tm_basemap("Esri.WorldImagery") +
tm_shape(non_overlapping) +
  tm_fill(pattern = "stripe", col="red", palette=rev(hcl.colors(4, "Purple-Blue")), style = "jenks", alpha=0.7, title="") +
tm_layout(legend.width=3, legend.text.size = .6, title.size = 1, legend.title.size = .1, asp=1.5, title = "Discontinuities Between Flood Hazard Identification Methods")

```



# Planned deviation for reproduction: Identify how many mobile homes would be likely to be in a River Corridor of the Connecticut River, if the creators of the River Corridor shapefile had included the Connecticut River.

This will help improve the geographic uncertainty caused by the boundary effect of the Connecticut River technically being fully part of New Hampshire, not Vermont.


```{r}
# Spatial join the buffered mobile home points to the estimated Connecticut River Corridor layer...not every MH will be within the estimated corridor
# Identifies which buffered mobile home points intersect the corridor
mh_ct_river_corridor_everything <- st_join(mh_buffer, ct_river_corridor, join = st_intersects)

# Filter to just the MHs that ended up in the flood zone...MHs in the river corridors meet the condition flood = TRUE
mh_ct_river_corridor <- mh_ct_river_corridor_everything %>%
  dplyr::filter(!is.na(WBNAME))

# Group by county, counting the number of MHs at risk
mh_ct_river_corridor_by_county <- mh_ct_river_corridor %>%
  group_by(county) %>%
  count(county, name = "mobile_home_count") %>%
  st_drop_geometry()

mh_ct_river_corridor_by_county
```

While this is definitely a bit of an overestimate, it does highlight how the VT River Corridor model's omission of the Connecticut River excludes a significant number of mobile homes.  There are a large number of mobile home parks close to the Connecticut River, which is a contributing factor.  61 mobile homes in Windsor at risk to flooding and 283 mobile homes in Windsor are at risk to flooding, aka they intersect my estimated river corridor for the Connecticut River.


To estimate a Connecticut River Corridor polygon layer (ct_river_corridor), I performed the following in QGIS:
- Load in a line layer for the centerline of the river channel, which I downloaded from the VT Open GeoData Portal
- I eyeballed the average width of the river channel by taking measurements with the measurement tool at various points along the river, and I made a well-informed assumption that the average channel width is 180 meters wide (this is probably a bit of an overestimate).  I then buffered the centerline by 90 meters to get a polygon that represents the river channel that is consistently 180 meters (buffered 90 meters from the centerline on either side).  
- I then added an additional 580 meter buffer.  This creates a polygon representing the river corridor that is consistently 1080 meters wide, or 6x the 180 meter average channel width

The documentation for the report on how River Corridors were constructed, on the Flood Ready Vermont website, said that a good rule of thumb is that a river corridor is 6x the width of the stream/river channel that it is based around.  While they do also take slope and geologic traits to provide a more specific river corridor profile, that is beyond the scope of this analysis, and I will make do with the 6x width approach.

## Check out the polygon that I made to estimate the River Corridor for the Connecticut River
```{r}
tmap_mode("view")
tm_basemap("Esri.WorldImagery") +
tm_shape(ct_river_corridor) +
  tm_fill(pattern = "stripe", col="lightblue", palette=rev(hcl.colors(4, "Purple-Blue")), style = "jenks", alpha=0.7, title="") +
tm_layout(legend.width=3, legend.text.size = .6, title.size = 1, legend.title.size = .1, asp=1.5, title = "Estimated River Corridor for the Connecticut River")
```


# Results

## Table 1: Mobile Homes at Risk by County (FEMA vs. River Corridors)

Ultimately, an R-based code approach to this spatial analysis study yields the exact same results that are found by Middlebury's GEOG 0261 class using the QGIS GUI.  I find that the number of total mobile homes (ACS data), mobile homes at risk (e911 and FEMA data), and mobile homes at risk (e911 and river corridor data) for Bennington, Rutland, Windham, and Windsor Counties are exactly the same. In Bennington County, the FEMA approach identifies significantly more mobile homes to be at risk than the River Corridor approach does.  In Windsor County, the two approaches predict about the same.  In Rutland and Windsor Counties, the River Corridor approach identifies significantly more mobile homes to be at risk than the FEMA approach.  These findings were true for both total number of mobile homes at risk **and** proportion of mobile homes in the county that are at risk.  


## Map 1 and Table 2: Percentage of Towns at Risk to Flooding (Either FEMA or River Corridor), by Town

A slight variation, as discussed earlier in the report, arises from a difference in the town shapefile layer.  The town layer that was provided to the GEOG 0261 class distinguishes between Rutland Town and Rutland City, while the town layer that I downloaded from the American Community Survey (since I was having corrupt file issues with the towns.shp used by the GEOG 0261 class) just shows Rutland as a single town.It is possible that the course instructors and I both started with the same town shapefile, but the instructors could have combined the two towns into one during their data processing.  Another reason could be that between the time the instructors downloaded the town shapefile and the time I downloaded the file, the town borders could have changed (ie. Rutland was broked into Rutland City and Rutland Town).  It is difficult to know exactly without knowing when the instructors downloaded their data and what they did to process it.


Reproducing the town-based risk map yielded the same exact results as the GEOG 0261 QGIS approach, except for Rutland.  When sorting towns based on highest percentage of their mobile homes at risk in the QGIS approach (Table 2), Rutland Town had the third highest % of mobile homes at risk out of all of the towns in the four counties.  However, when considering Rutland as a single entity in my R reproduction, it does make the top 10 list.  Instead, all the towns get shifted up one place, with Proctor, VT being the town with the 10th highest risk.  There are some threats to validity here, namely the small number problem, which I will addreess in the Discussion section.

Woodford and Woodstock have the first and second highest % of mobile homes at risk in both R and QGIS, with 85.18% and 68.57% respectively of their mobile homes at some risk to flooding, as determined by FEMA or River Corridors




## R code vs. QGIS GUI: 

Ultimately, the main goal of this reproduction was to find a way using R code to do all the spatial analysis steps that are taught to GEOG 0261 students in QGIS in the "Flood Hazard Vulnerability in Vermont's Mobile Homes" assignment for the course.  I followed the exact same workflow provided by the instructors of GEOG 0261 to create Table 1.  While I technically could have done a select by location to identify the mobile homes in the flood zones and river corridors, I chose to use a join by location and then a filter because of a warning that was thrown and because it was computationally faster.


I utilized the stars and tmap libraries to work with the spatial data.





## Addition of the estimated Connecticut River corridor: 

Based on my estimation of the river corridor for the Connecticut River, I found that 61 mobile homes in Windham and 283 in Windsor are missed when using the existing river corridor shapefile that is published on the VT Open GeoData Portal and is provided to the GEOG 0261 students.  Although I likely overestimate the extent of the river corridor a bit (since I did not account for geologic conditions or topography/slope...I only considered meander belt width), this shows that VT River Corridors do miss out on identifying mobile homes that could be at risk.  Fortunately, the FEMA Flood Zones do seem to contain most of the mobile homes along the Connecticut River that do appear to be at risk (see the presence of FEMA Flood Zones in the first map presented at the top of this report...the map of Flood Zones).




# Discussion


## Sources of uncertainty and geographic threats to validity:
Construct validity
Small number problem if a town doesn't have a ton of mobile homes...for rates
Lineage section to specify the data sources


### Future additions to the study


## Pedagogical implications: 


## Policy implications: 


### Connecticut River:





# Conclusion





























## Analysis

Describe the methods of analysis that will directly test the hypotheses or provide results to answer the research questions.
This section should explicitly define any spatial / statistical *models* and their *parameters*, including *grouping* criteria, *weighting* criteria, and *significance thresholds*.
Also explain any follow-up analyses or validations.



# Integrity Statement
I followed and accomplished what I set out to do in my preanalysis plan.  I did make a few slight deviations to improve the code and the interpretability of my report/findings:
- To isolate the mobile homes in the flood risk polygons, instead of selecting by location (1 step), I ended up using a spatial join and then filtering based on mobile homes that were in the flood risk layer(2 steps) because of a warning message I got with the 1 step approach.  The 1 step approach took much longer to run.
- I ended up creating a polygon layer that is the difference between FEMA 100yr Flood Zones and VT River Corridor polygons, to show where the two methods of spatially identifying flood risk do not line up.  I then plotted this in an interactive tmap with a satellite basemap so that the reader can play around with it to see where they don't align.

I - the author of this reproduction study - state that I completed what I outlined in my preregistration to the best of my knowledge and that no other preregistration exists pertaining to the same hypotheses and research.

# Acknowledgements

This report is based upon the template for Reproducible and Replicable Research in Human-Environment and Geographical Sciences, DOI:[10.17605/OSF.IO/W29MQ](https://doi.org/10.17605/OSF.IO/W29MQ)

# References
Baker, D., Hamshaw, S. D., & Hamshaw, K. A. (2014). Rapid flood exposure assessment of Vermont mobile home parks following Tropical Storm Irene. Natural Hazards Review, 15(1), 27-37. DOI: 10.1061/(ASCE)NH.1527-6996.0000112.

Schmitt, R. R. (1978). Threats to validity involving geographic space. Socio-Economic Planning Sciences, 12(4), 191–195. https://doi.org/10.1016/0038-0121(78)90044-7